---
title: "Gap Spacing in CA NEON Sites"
author: "Jon Weiner"
date: "June 22, 2016"
output: html_document
---

The goal of this project is to try and determine gap spacing across the CA NEON sites (TEAK, SOAP, and SJER). We will focus on the NEON plots that are in each of those areas.

#### Initialize Libraries

```{r import_libraries}
#import libraries
library(raster)
library(rgdal)

```

#### Load CHM from TEAK, SOAP, and SJER and do initial plots

```{r load_chms}
#load CHMs

SJER_chm <- raster("../NEONdata/D17-California/SJER/2013/lidar/SJER_lidarCHM.tif")
SJER_chm

plot(SJER_chm,
     main = "SJER CHM")

SOAP_chm <- raster("../NEONdata/D17-California/SOAP/2013/lidar/SOAP_lidarCHM.tif")
SOAP_chm

plot(SOAP_chm,
     main = "SOAP CHM")

TEAK_chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")
TEAK_chm

plot(TEAK_chm,
     main = "TEAK CHM")

#plot histograms
hist(log(SJER_chm),
     main = "SJER CHM Hist in Log")
hist(log(SOAP_chm),
     main = "SOAP CHM Hist in Log")
hist(log(TEAK_chm),
     main = "TEAK CHM Hist in Log")
```

#### Open random locations

```{r random_locations}

SJER_rand_locs <- sampleRandom(SJER_chm, size = 100, xy = TRUE, sp = TRUE)
SOAP_rand_locs <- sampleRandom(SOAP_chm, size = 100, xy = TRUE, sp = TRUE)
TEAK_rand_locs <- sampleRandom(TEAK_chm, size = 100, xy = TRUE, sp = TRUE)

```


#### Test plot locations for sanity sake
```{r plot_locations}

plot(SJER_chm)
plot(SJER_rand_locs,
     add = TRUE)

plot(SOAP_chm)
plot(SOAP_rand_locs,
     add = TRUE)

plot(TEAK_chm)
plot(TEAK_rand_locs,
     add = TRUE)

```

### Extract values from points and calculate SD for SJER

```{r extract_raster_info_SJER}
#Extract values from points and calculate SD for SJER

# 10m BUFFER

SJER_locs_with_sd_10m <- extract(SJER_chm,
                       SJER_rand_locs,
                       buffer = 10,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(SJER_locs_with_sd_10m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
SJER_sd_10m <- SJER_locs_with_sd_10m@data[,4]

#Plot histogram
hist(SJER_sd_10m,
     main = "SJER_10m_SD",
     xlab = "Standard Deviation")

# 25m BUFFER

#SJER
SJER_locs_with_sd_25m <- extract(SJER_chm,
                       SJER_rand_locs,
                       buffer = 25,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(SJER_locs_with_sd_25m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
SJER_sd_25m <- SJER_locs_with_sd_25m@data[,4]

#Plot histogram
hist(SJER_sd_25m,
     main = "SJER_25m_SD",
     xlab = "Standard Deviation")

# 50m BUFFER

#SJER
SJER_locs_with_sd_50m <- extract(SJER_chm,
                       SJER_rand_locs,
                       buffer = 50,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(SJER_locs_with_sd_50m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
SJER_sd_50m <- SJER_locs_with_sd_50m@data[,4]

#Plot histogram
hist(SJER_sd_50m,
     main = "SJER_50m_SD",
     xlab = "Standard Deviation")

# 100m BUFFER

#SJER
SJER_locs_with_sd_100m <- extract(SJER_chm,
                       SJER_rand_locs,
                       buffer = 100,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(SJER_locs_with_sd_100m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
SJER_sd_100m <- SJER_locs_with_sd_100m@data[,4]

#Plot histogram
hist(SJER_sd_100m,
     main = "SJER_100m_SD",
     xlab = "Standard Deviation")

```

### Extract values from points and calculate SD for SOAP

```{r extract_raster_info_SOAP}
#Extract values from points and calculate SD for SOAP

# 10m BUFFER

SOAP_locs_with_sd_10m <- extract(SOAP_chm,
                       SOAP_rand_locs,
                       buffer = 10,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(SOAP_locs_with_sd_10m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
SOAP_sd_10m <- SOAP_locs_with_sd_10m@data[,4]

#Plot histogram
hist(SOAP_sd_10m,
     main = "SOAP_10m_SD",
     xlab = "Standard Deviation")

# 25m BUFFER

SOAP_locs_with_sd_25m <- extract(SOAP_chm,
                       SOAP_rand_locs,
                       buffer = 25,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(SOAP_locs_with_sd_25m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
SOAP_sd_25m <- SOAP_locs_with_sd_25m@data[,4]

#Plot histogram
hist(SOAP_sd_25m,
     main = "SOAP_25m_SD",
     xlab = "Standard Deviation")

# 50m BUFFER

SOAP_locs_with_sd_50m <- extract(SOAP_chm,
                       SOAP_rand_locs,
                       buffer = 50,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(SOAP_locs_with_sd_50m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
SOAP_sd_50m <- SOAP_locs_with_sd_50m@data[,4]

#Plot histogram
hist(SOAP_sd_50m,
     main = "SOAP_50m_SD",
     xlab = "Standard Deviation")

# 100m BUFFER

SOAP_locs_with_sd_100m <- extract(SOAP_chm,
                       SOAP_rand_locs,
                       buffer = 100,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(SOAP_locs_with_sd_100m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
SOAP_sd_100m <- SOAP_locs_with_sd_100m@data[,4]

#Plot histogram
hist(SOAP_sd_100m,
     main = "SOAP_100m_SD",
     xlab = "Standard Deviation")

```

### Extract values from points and calculate SD for TEAK

```{r extract_raster_info_TEAK}

#Extract values from points and calculate SD for TEAK

# 10m BUFFER

TEAK_locs_with_sd_10m <- extract(TEAK_chm,
                       TEAK_rand_locs,
                       buffer = 10,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(TEAK_locs_with_sd_10m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
TEAK_sd_10m <- TEAK_locs_with_sd_10m@data[,4]

#Plot histogram
hist(TEAK_sd_10m,
     main = "TEAK_10m_SD",
     xlab = "Standard Deviation")

# 25m BUFFER

TEAK_locs_with_sd_25m <- extract(TEAK_chm,
                       TEAK_rand_locs,
                       buffer = 25,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(TEAK_locs_with_sd_25m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
TEAK_sd_25m <- TEAK_locs_with_sd_25m@data[,4]

#Plot histogram
hist(TEAK_sd_25m,
     main = "TEAK_25m_SD",
     xlab = "Standard Deviation")

# 50m BUFFER

TEAK_locs_with_sd_50m <- extract(TEAK_chm,
                       TEAK_rand_locs,
                       buffer = 50,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(TEAK_locs_with_sd_50m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
TEAK_sd_50m <- TEAK_locs_with_sd_50m@data[,4]

#Plot histogram
hist(TEAK_sd_50m,
     main = "TEAK_50m_SD",
     xlab = "Standard Deviation")

TEAK_locs_with_sd_100m <- extract(TEAK_chm,
                       TEAK_rand_locs,
                       buffer = 100,
                       fun = sd,
                       sp = TRUE,
                       stringAsFactors = F)
#Rename Columns
names(TEAK_locs_with_sd_100m@data) <- c("x", "y", "Loc_Ht", "Std Dev" )

#Pull Std Devs
TEAK_sd_100m <- TEAK_locs_with_sd_100m@data[,4]

#Plot histogram
hist(TEAK_sd_100m,
     main = "TEAK_100m_SD",
     xlab = "Standard Deviation")

```



